---
title: "Clinical Portfolio"
author: "Stephanie B. Ward, M.S."
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: true
    code-summary: "Hide Code"
editor:
  chunk-output-type: inline
---

```{css}
pre, code {
  max-height: 500px;
  overflow-y: auto;
  white-space: pre !important; 
  overflow-x: auto
}
```

```{r}
#| message: false
#| echo: false

# Packages and Source
library(tidyverse)
library(psych)
library(janitor)
library(jmv)
library(kableExtra)
library(lubridate)
```

```{r}
#| message: false
#| echo: false

# Global Settings
theme_set(theme_classic())
options(tibble.width = Inf)
options(tibble.print_max = Inf)
```

## Background

### Personal Statement

I collect moments that clarify my purpose. Among the most powerful are those shared with patients grappling with the questions my research seeks to answer. One such moment belonged to a college student struggling to determine whether an intoxicated sexual experience was coercive or consensual. Alcohol-related sexual violence is the focus of my work, and even then, I could not protect this young woman from what had already happened. Her session laid bare the blurry lines available for navigating ambiguous sexual events and their consequences.

My long-term career goal is to work as part of an interdisciplinary research team focused on promoting women's health by reducing interpersonal violence and hazardous substance use. Specifically, I want to leverage mobile technologies and the scaffolding of preexisting systems to deliver scalable, evidence-based interventions in clinical and educational settings. I am committed to disentangling complexities across the spectrum of sexual behavior, from consensual to non-consensual and sober to incapacitated via substance use, because I believe facilitating a more comprehensive understanding of these behaviors can transform the sexual violence prevention landscape.

A paradigm shift is sorely needed given the current state of the primary prevention literature and I am on my way to becoming a well-rounded psychological scientist who advances systematic change. In this vein, I have deliberately integrated my teaching, research, and clinical training experiences to prepare for a dynamic career that makes the best use of my expertise across these domains. For example, I led discussion sections on human sexuality to further my goal of using comprehensive sexuality education to prevent sexual violence among college students, then went on to complete a clinical practicum rotation at our institution's health center to inform my research focused on this population.

Notably, my path has not been without challenges. I was the first student in a new lab established during a global pandemic and then diagnosed with a rare genetic condition that causes chronic pain. My illness tested my resilience as it forced me to balance the demands of a brilliantly rigorous graduate program while fighting to be taken seriously in our healthcare system. Perseverance was critical because for me and countless others, seeking treatment means signing up to be judged and patronized with your pain dismissed. The emotional consequences of these interactions are also common among victims of sexual violence: feeling powerless, confused, and alone.

Experiencing this powerlessness firsthand clarified my purpose on a deeper level. My research aims to help stop sexual violence before it starts because I know what it feels like when your bodily autonomy is denied and will never be satisfied dealing solely with the fallout. I provide trauma-informed care because patients deserve to be seen and heard--especially by those advancing science via knowledge derived from their most vulnerable moments. I am honored to do this work and look forward to tackling new challenges in the years ahead, secure in the knowledge that with each passing milestone I get one step closer to enacting positive change for all the patients whose moments I carry with me.

### Curriculum Vitae

CV can be accessed here

### Transcript

Transcript can be accessed here

## Research Experiences

### Research Statement

My research program aims to reduce sexual violence via a deeper understanding of sexual behavior following alcohol consumption. In pursuit of this goal, I integrate qualitative, quantitative, and experiencing sampling methods to investigate how processes spanning multiple ecological levels shape sexual consent dynamics and patterns of substance use, thereby heightening the risk for sexual violence.

Personal characteristics, relationships, and broader social forces shape the lens through which we interpret our experiences. The information we use to form belief systems and make decisions is filtered through this lens, which is highly consequential in the context of sexual behavior. For instance, the #MeToo movement drew the public's attention to sexual violence and changed how some adults view consent and approach sexual interactions (Ward, Nardella, Hamilton & Walsh, 2024). Awareness of #MeToo and the nature of resultant changes varied as a function of political ideology and gender, demonstrating how individual- and societal-level factors interactively influence sexual consent.

In our #MeToo study, qualitative responses identified sexual victimization history and substance use as issues that complicate sexual consent, which led us to conduct subsequent research examining how these individual-difference factors are associated with sexual consent feelings and behaviors. Histories of sexual violence victimization were indirectly related to the use of nonverbal and passive sexual consent cues through lower consent feelings (Walsh, Lowe, & Ward, 2024), suggesting that mutual consideration and communication throughout sexual events may help to ensure both parties are willingly participating. Alcohol may undermine this ongoing process, so I led a study examining attitudes toward alcohol-involved sexual experiences and process-based consent among college men. Our findings revealed a significant three-way interaction indicating a negative association between positive attitudes toward alcohol-involved sexual experiences and process-based consent, which was exacerbated when both hostility toward women and peer support of abuse were high (Ward, Hassanpour, & Walsh, under review). Collectively, these findings clarify the interplay between intrapersonal, interpersonal, and cultural risk factors that may give rise to substance-facilitated sexual violence.

My future research will progress the science in two important ways. Traditional quantitative approaches can miss the detailed insights needed to contextualize study findings while behavioral measures of substance use and consent often rely on retrospective self-reports. To gain a more holistic perspective of the proximal and temporal associations between sexual behavior and substance use, I will analyze qualitative and longitudinal quantitative data in my dissertation. Specifically, I will use qualitative methods to explore how young adults navigate intoxicated sexual experiences and determine for themselves whether these experiences are consensual. My lab and I also collected daily diary data from college students to investigate how substance use and subjective intoxication influence sexual risk and consent dynamics at the event level. Experience sampling methods are acceptable to people who use substances (Wyant, Moshontz, Ward, Fronk, & Curtin, 2023), and surveying students 'in the wild' can address methodological hurdles by enabling us to probe event-level associations. Collecting rich qualitative data and dynamic daily reports will strengthen my examination of intoxicated, consent-related behaviors by facilitating a more nuanced analysis of variability within and between events.

My ultimate program of research will leverage the strengths of multimodal ambulatory assessment and mixed methodologies to better account for the real-world complexities underlying what happens during and after substance-involved sexual experiences. I intend to advance our field's understanding of the precursors and contextual factors that contribute to these incidents on college campuses and then expand the scope of my work to include members of the armed forces and individuals impacted by the criminal legal system. My primary goal is to inform, develop, and deliver risk-reduction and health promotion interventions aimed at preventing sexual violence and fostering healthy, consensual sexual behaviors in contexts where alcohol and other substances are involved.

### Research Products

#### Submitted Version of the FYP

1.  **Ward, S. B.**, Hassanpour, P., Walsh, K. (2023). Process-based consent among college men: The interactive influence of hostility toward women, perceived peer norms, and alcohol-involved sexual consent attitudes. *Sex Roles.* Registration DOI (anonymized view-only link for masked peer review): <https://osf.io/zxrjp/?view_only=89295c61cb5f415c9f9526df9352d517> 

Relative contributions:

-   Conceptualization - 60%
-   Design - 60%
-   Analyses - 90%
-   Writing - 70%

**Abstract**

Process-based consent refers to the ongoing consideration of a partner's willingness to engage in sexual activity and continual communication with that partner. Drawing upon research identifying individual-, situational-, and peer-level factors that contribute to men's perpetration of sexual violence, the current study used online survey data and hierarchical linear regression to examine how permissive attitudes toward sexual consent experiences involving alcohol use, hostility toward women, and perceived peer norms in support of abuse influence process-based consent among college men (N = 463; M age = 19.32; 93% heterosexual, 64% white, non-Hispanic). Findings revealed all three primary study variables were independently associated with weaker process-based consent while adjusting for fraternity membership, relationship status, and sexual history. A significant three-way interaction indicated that the negative association between permissive alcohol-involved sexual consent attitudes and process-based consent was exacerbated when both hostility toward women and peer support of abuse were high, implying a complex relationship between multiple forces shaping college men's approaches to sexual consent. Implications for consent education and violence prevention are discussed, including the need for interventions that address process-level interactions between personal beliefs and gendered peer group norms concerning alcohol use and sexual behavior.

#### Published Empirical Studies

1.  **Ward, S. B.**, Nardella, S., Hamilton, K. W., Walsh, K. (2024). "I Didn't Realize How Common It Was:" A Mixed-Methods Study Examining Changes in Perceptions of Sexual Assault, Sex and Consent, and Sexual Behavior as a Function of the #MeToo Movement. *The Journal of Sex Research*, 1--15. <https://doi.org/10.1080/00224499.2024.2352555>

Relative contributions:

-   Conceptualization - 50%
-   Design - 50%
-   Analysis - 20%
-   Writing - 50%

**Abstract**

This mixed-methods study examined awareness and perceived legitimacy of the #MeToo movement and how #MeToo changed perceptions of sexual assault and consent, as well as sexual interactions, in the United States. Adults residing in the U.S. were recruited through CloudResearch to complete an online survey in 2021. Quantitative data from 680 participants (*M* age = 45.8, 60% women, 77.4% White) indicated moderate awareness and perceived legitimacy of the #MeToo movement; Black, LGBQ+, and more politically liberal respondents had greater #MeToo awareness while younger, more liberal respondents, and those with greater rejection of rape myths rated #MeToo more legitimate. Among 354 participants (*M* age = 45.0, 65.3% women, 76.4% White) who answered at least one open-ended question with regard to changes resulting from the #MeToo movement, thematic analysis revealed nine primary themes: (1) Describing change; (2) Change in understanding; (3) More cautious; (4) Wrong or requires reporting/punishment; (5) Gendered social norms; (6) Easier to talk about; (7) #MeToo rhetoric; (8) Consequences for survivors, and (9) Empowerment. Several sub-themes were identified among the primary themes and implications for sexual assault prevention and response were discussed. The study findings are contextualized by social constructionism, with a particular focus on gender-based power dynamics, contributing to a growing literature documenting the cultural impact of the #MeToo movement.

2.  Walsh, K., Lowe, S. R., & **Ward, S.** (2024). Sexual Victimization, Emotion Dysregulation, and Sexual Consent Feelings and Communication During a Recent Sexual Encounter. *The Journal of Sex Research*, 1--9. <https://doi.org/10.1080/00224499.2024.2368044>

Relative contributions:

-   Conceptualization - 20%
-   Design - 20%
-   Analysis - NA
-   Writing - 30%

**Abstract**

Sexual consent has been a major focus of campus campaigns to reduce sexual violence (SV). However, these campaigns often educate students about consent with little attention to the complex ways consent can be experienced, expressed, and interpreted by others. Further, little research has focused on the consent feelings and communication cues of students who have a history of SV, nor have studies examined how the ability to attend to and regulate emotions relates to internal feelings or external communication of consent. This secondary analysis examined SV histories, emotion dysregulation, and internal and external consent cues in a sample of 610 college students (72% women) who completed an online self-report survey. Findings revealed that students with SV histories (n = 257) reported greater emotion dysregulation, lower internal consent, and greater use of passive external consent cues compared to students without SV histories; greater emotion dysregulation was associated with lower internal consent. In a path model, SV history was indirectly related to use of direct nonverbal and passive consent cues through greater emotion dysregulation and lower internal consent. Contrary to expectations, higher internal consent was associated with greater use of passive external consent cues. Consent programming could be augmented to encourage attention to and regulation of emotions in sexual situations to improve communication and sexual well-being.

#### Cross-Lab Collaborations

1.  Wyant, K., Moshontz, H., **Ward, S.B.**, Fronk, G., Curtin, J. J. (2023). Acceptability of Personal Sensing Among People With Alcohol Use Disorder: Observational Study. *JMIR Mhealth Uhealth*, 11(e41833), <https://doi.org/10.2196/41833>.

Relative contributions:

-   Conceptualization - 10%
-   Design - 10%
-   Analysis - NA
-   Writing - 10%

#### Post-bacc Publications

1.  **Ward, S. B.**, Boger, M., Fleishman, A., Shenkel, J., Calvo, A., Pomahac, B., Zwolak, R., Krishnan, N., Richman, K., Rodrigue, J.R. (2020). Attitudes toward organ, tissue and vascularized composite allograft (VCA) donation and transplantation: A survey of United States military veterans. Transplantation. <https://doi.org/10.1097/tp.0000000000003376>

2.  Robinson, K. A., Carroll, M., **Ward, S. B.**, Osman, S., Chhabra, K. R., Arinze, N., Amedi, A., Kaafarani, H., Smink, D., Kent, T., Musa M. Aner, Brat, G. (2020). Implementing and evaluating a multihospital standardized opioid curriculum for surgical providers. Journal of Surgical Education, 77(3), 621--626. <https://doi.org/10.1016/j.jsurg.2019.12.012>

#### Capstone Presentation

*Sexual Violence in University Students: A Multi-Level Approach to Understanding the Perpetration of Sexual Aggression*

-   Date presented: April 26, 2024

**Abstract**

#### Lunch and Learn Presentation

*Process-Based Consent Among College Men: The Interactive Influence of Hostility Toward Women, Perceived Peer Norms, and Alcohol-involved Sexual Consent Attitudes*

-   Date Presented: February 20, 2023  

**Abstract**

Process-based consent refers to the ongoing consideration of a partner's willingness to engage in sexual activity and continual communication with that partner. Drawing upon research identifying individual-, situational-, and peer-level factors that contribute to men's perpetration of sexual violence, the current study used online survey data and hierarchical linear regression to examine how permissive attitudes toward sexual consent experiences involving alcohol use, hostility toward women, and perceived peer norms in support of abuse influence process-based consent among college men (N = 463; M age = 19.32; 93% heterosexual, 64% white, non-Hispanic). Findings revealed all three primary study variables were independently associated with weaker process-based consent while adjusting for fraternity membership, relationship status, and sexual history. A significant three-way interaction indicated that the negative association between permissive alcohol-involved sexual consent attitudes and process-based consent was exacerbated when both hostility toward women and peer support of abuse were high, implying a complex relationship between multiple forces shaping college men's approaches to sexual consent. Implications for consent education and violence prevention are discussed, including the need for interventions that address process-level interactions between personal beliefs and gendered peer group norms concerning alcohol use and sexual behavior.

#### Posters

*\*\*Mentored presentation by an undergraduate mentee.*

1.  \*\*Does Anxiety Moderate the Effect of Peer Drinking Norms on Past-Month Alcohol-Induced Blackouts? An Observational Study Among College Students.\

2.  \*\*Associations Between Lifetime Sexual Violence and Engaging in Choking During Sex: Findings from an Undergraduate Sample.\

3.  \*\*Manner of Drinking as a Mediator of Emotional Impulsivity and Alcohol-Related Problems.\

4.  Alcohol-Induced Blackouts and Nonconsensual Sexual Experiences Among Undergraduate Men.\

5.  Sexual Consent Among College Men: The Moderating Effects of Alcohol-Involved Sex and Hostility Toward Women.\

6.  Associations between Sexual Assault History and Sexual Consent Feelings and Behaviors among College Students: A Path Model Involving Alcohol Use Prior to Sex & Emotion Dysregulation.\

7.  \*\*Sexual Assault Resistance Strategies, Blame, and Mental Health Outcomes.\

8.  Sexual Assault Victimization and Perpetration among those Involved in Greek Life and Varsity Athletics: A Large National Study.\

9.  Multi-Institutional Patient-Reported Opioid Use Compared to Clinical Consensus Recommendations.\

10. Comparing Reality to Consensus: Reported Opioid Use vs. Clinician Consensus Recommendations.

#### Workshops

**Attended**

-   Academic Job Materials Workshop - Princeton Psychology Postdoctoral DEI Committee
-   Mixture Modeling and Latent Class Analysis -- Daniel J. Bauer, PhD, and Doug Steinley, PhD
-   Responsible Authorship and Publication -- Margarita Kaushanskaya, PhD,
-   Research Misconduct and Policies for Handling Misconduct -- Brain Fox, PhD
-   Ins and Outs of F and K Awards -- Waisman Center PIs and Grants Team
-   Data Acquisition, Analysis, Sharing, and Ownership -- Brian Christian, PhD
-   Collaborative Research -- Doug Dean III, PhD
-   Mentor and Mentee Responsibilities and Relationships -- Rebecca Alper, PhD, CCP‐SLP
-   Peer Reviews and NIH Grant Reviews -- Audra Sterling, PhD
-   Conflict of interest---Personal, Professional, and Financial -- Maureen Durkin, PhD, DrPH
-   Ethical Conduct of Human and Animal Subjects Research -- Katherine Hustad, PhD, CCC‐SLP
-   Secure and Ethical Data Use and Management -- Brittany Travers, PhD
-   Introduction to Structural Equation Modeling -- Daniel J. Bauer, PhD, and Patrick J. Curran, PhD

**Focused on DEI Practices and Principles**

-   Breaking the Bias Habit -- William T.L. Cox, PhD
-   Promoting Safe and Inclusive Research Environments -- Leann DaWalt, PhD
-   The Scientist as a Responsible Member of Society -- Karla Ausderau, PhD
-   Change in Implicit Bias -- Calvin Lai, PhD
-   Cross-Race Friendship -- Elizabeth Page-Gould, PhD
-   Exploring 'the story' of inequality: Scientific investigations into the antecedents, experience, and reduction of group-based inequality -- Demis Glasford, PhD
-   Inequalities in Young Adult Criminal Justice Contacts: Patterns and Collateral Consequences for Parents and Siblings -- Courtney Boen, PhD, MPH

**Led**

Described below in the Teaching and Mentorship section.

## Clinical Experiences

### Theoretical Orientation Statement

### Clinical Training Record

#### Practicum Experiences

[*Brief description should include the name and dates for the practicum, brief description of the client population and other relevant details (e.g., interventions, modalities). This should include documentation of clinical hours (per internship categories) and available/completed supervisor evaluations.*]{.underline}

**University Health Services, University of Wisconsin-Madison**

September 1, 2022 - May 13, 2023, 16 - 20 hours per week

-   Individual Counseling: 5 -- 8 hours/week.
-   Group Counseling: 2 hours/week.
-   Individual Supervision: 1.5 hours/week.
-   Group Supervision: 1.5 hours/week.
-   In-service Training: 1.5 hours every other Friday.
-   Group Therapy Seminar/Supervision: 1.5 hours every other Friday.
-   Administrative/Case Notes: 3-5 hours/week.

As a practicum trainee in the Mental Health Services division of our institution's health center, I maintained a clinical caseload of 10-18 young adults. My patients were undergraduate and graduate students who presented with symptoms consistent with social and generalized anxiety, posttraumatic stress, attention-deficit/hyperactivity disorder, and substance dependence, though concerns specific to depression, suicidal thoughts, self-injurious behaviors, poor body image/disordered eating and low self-worth were also quite common. During this clinical practicum, I gained experience working as a multidisciplinary team member within a coordinated care model. I collaborated with clinicians from other service divisions including crisis stabilization, primary care, OB/GYN, and psychiatry to provide holistic care to each patient. While at UHS and thereafter, I drew most heavily from third-wave cognitive behavioral approaches including dialectical behavior therapy (DBT), acceptance and commitment therapy (ACT), and motivational interviewing. In facilitating "DBT Skills" and "ACT for Social Anxiety" therapy groups, I deeply engaged with these modalities and used what I learned to cultivate a more grounded professional identity. Through this practicum experience, my therapeutic approach grew to center mindfulness, values, and emotions, employing culturally-informed conceptualization and goal-oriented treatment planning.

```{r,echo=FALSE}
#| label: tbl-uhs
#| tbl-cap: "Hours by Type at University Health Services"

# create data frame
uhs_summ <- data.frame(
  Summary = c("Individual Therapy", "Group Counseling", "Assessment", "Total Direct", "Total Support"),
  Hours = c("117", "36", "27", "180", "87.25"))
# format table
kable(uhs_summ, booktabs = TRUE, align = 'lr', col.names = c("Summary", "Hours")) %>%
  kable_styling(full_width = FALSE)
```

#### Integrative Assessment Reports

De-identified assessment reports are available here.

1\. PDD

2\. ADHD

3\. ASPD

4\. ED - BN

5\. cPTSD, cross-listed with DEI

## Diversity, Equity, and Inclusion

### Diversity Statement

Systemic issues that threaten diversity, equity, and inclusion are inextricably linked to the roles I inhabit as a feminist research scientist and aspiring clinical psychologist.

I study sexual violence, the burden of which is inequitably distributed. Racialized and other minoritized groups are disproportionately impacted by sexual violence and the risk of victimization is compounded among those with multiple marginalized identities. I promote equity through sexual violence research by investigating the nature of these disparities and facilitating efforts to eliminate them. Producing rigorous, high-quality science in this area of study requires cultural humility, awareness of subjectivity, and recognition of racist, patriarchal power structures.

To this end, I subscribe to the contemporary constructivist philosophy of science and use theory aligned with its core assumptions to guide my research. For example, a constructionist framework contradicts the positivist notion of a mind-independent, absolute truth, instead positing that reality is malleable and "constructed" through social processes (DeLamater & Hyde, 1998). This framework assumes social problems are actively made, not simply revealed, a core tenet of feminist science and the scholarly literature on societal norms that perpetuate sexual and gender-based violence.

Social problems are actively made in society at large but also within our communities and institutions. Promoting diversity, equity, and inclusion therefore requires turning inward to examine the principles and practices used in the conduct of our research. There are current practices in psychological science that exclude scholars from marginalized backgrounds, which perpetuates white supremacy (Buchanan et al., 2021). I use open science practices that facilitate sharing and reuse of research assets to help mitigate inequalities disadvantaging under-resourced scholars via increased accessibility. The open science movement explicitly features collaborative, prosocial ideals known to engage members of under-represented groups and cultivate a more inclusive culture (Murphy et al., 2020). These ideals reflect my values and align with my long-term goal of working as part of a cooperative research team.

### DEI Training Record

#### Case Conceptualization

The integrated intake assessment, conceptualization, treatment plan, and case summary for a patient with intersecting marginalized identities is accessible here.

#### DEI - Workshops, Seminars, and Lectures

**Research**

-   Breaking the Bias Habit -- William T.L. Cox, PhD
-   Promoting Safe and Inclusive Research Environments -- Leann DaWalt, PhD
-   The Scientist as a Responsible Member of Society -- Karla Ausderau, PhD
-   Change in Implicit Bias -- Calvin Lai, PhD
-   Cross-Race Friendship -- Elizabeth Page-Gould, PhD
-   Exploring 'the story' of inequality: Scientific investigations into the antecedents, experience, and reduction of group-based inequality -- Demis Glasford, PhD
-   Inequalities in Young Adult Criminal Justice Contacts: Patterns and Collateral Consequences for Parents and Siblings -- Courtney Boen, PhD, MPH

**Clinical Service Provision**

-   International Student-Clients: Clinical Considerations -- Wei-Chiao Hsu 徐薇喬, PhD, LPC
-   Recognizing and Addressing Bicultural Identity Conflict -- Geetanjali Deole, MS, LPC

#### DEI - Mentorship

During the summer of 2021, I had the opportunity to mentor a promising undergraduate student, Rachna Iyer, through the Psychology Research Experience Program (PREP) at UW-Madison. This program, supported by the National Science Foundation, focuses on providing students from underrepresented backgrounds with intensive research experiences in psychology and neuroscience.

In my role as a mentor, I introduced Rachna to the sexual violence literature pertaining to posttraumatic stress, providing her with a foundational understanding necessary for conducting meaningful research in her area of interest. Over the course of the program, I guided her through the entire scientific process---from the initial stages of identifying a research question to the final stages of writing up the results.

One of my primary goals was to equip Rachna with practical skills in data analysis and management. I facilitated her learning of analytic strategies using R and RStudio, which are critical tools in psychological research. This training not only enhanced her technical skills but also her confidence in handling complex datasets.

Additionally, I supported Rachna in her professional development by helping her prepare for presentations. She successfully presented her project at the 2021 PREP Symposium and the International Society for Traumatic Stress Studies (ISTSS) conference, opportunities that allowed her to engage with the broader scientific community and gain valuable feedback on her work.

This mentorship experience was deeply enriching, as it allowed me to contribute to the growth of an aspiring clinical scientist from an under-represented background in academic psychology. After the PREP program, Rachna graduated early from the University of Michigan and then worked for Dr. Nicole Weiss, an Associate Professor and Clinical Psychologust at the University of Rhode Island, whose research program focuses on the co-occurrence of PTSD and substance use disorders.

## Teaching and Mentorship

### Teaching Experiences

[*Include course evaluation summary statistics and raw evaluations.*]{.underline}

Starting in the Spring semester of 2022, I served as Dr. Janet Hyde's Head Teaching Assistant for the Psychology of Human Sexuality course (PSYC/SOC 453). I led four weekly discussion sections, each consisting of 15-20 undergraduates, encouraging critical thinking and deeper engagement with course materials. My responsibilities extended to coordinating testing, managing section changes, and overseeing the grading process for the 300-student lecture course. I also delivered two guest lectures on love and gender, which were well-received by students.

During the Fall semester of 2022, I was a Teaching Assistant for the Introductory Psychology course (PSYC 202), taught by Dr. Rebecca Addington. This large, 400-student lecture course required meticulous organization and pedagogical skills; I was responsible for grading approximately 400 short essay assignments each week. Additionally, I managed logistical aspects of the course such as scheduling, proctoring, and hand-scoring makeup exams. I also supported students through tutoring during office hours, by appointment, and via email, ensuring their academic and personal support.

In January, 2024, I delivered a Guest Lecture at the School of Pharmacy for Karen Kopacek, MS, RPh. The lecture focused on emotional stress and compassion fatigue and was psychoeducational in nature, as I integrated practical skills such as mindfulness, distress tolerance, and emotion regulation to support future healthcare professionals.

These experiences have shaped my teaching philosophy and approach, reinforcing my commitment to fostering an interactive and supportive educational environment.

### Mentorship Experiences

Since August 2020, I have been actively involved in mentorship in the Preventing Interpersonal Violence & Overcoming Trauma (PIVOT) Lab at UW-Madison. Under the direction of my advisor, Dr. Kate. Walsh, our lab's mission is to address critical issues related to sexual violence and trauma through rigorous research and scholarly engagement.

I have provided mentorship to a diverse group of lab members, ranging from undergraduates to post-baccalaureate and master's level research assistants. My contributions have focused on enhancing their research competencies, including data management and cleaning using RStudio.

Additionally, I have led workshops aimed at enhancing research proficiency, covering topics such as effective literature search strategies and the use of Zotero for organizing and citing scholarly articles. These workshops are designed to enhance lab members' practical, hands-on experience and are tailored to meet the evolving needs of our research program.

A significant part of my mentorship experience has also been devoted to guiding our lab members through the graduate application process. I provide tailored support to help each individual identify and apply to graduate programs and paid training opportunities that match their career aspirations in research and clinical practice. This guidance is customized to reflect the unique goals and backgrounds of each mentee, ensuring that they receive the most relevant and impactful advice.

Through these activities, the PIVOT Lab advances research and promotes professional development among emerging scholars in psychology and related fields.
