---
title: Clinical Portfolio
date: 2024-07-26
author:
  - name: Stephanie B. Ward, M.S.
    affiliation: University of Wisconsin-Madison
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: true
    code-summary: "Hide Code"
    execute:
      echo: false
editor:
  chunk-output-type: inline
---

```{css}
pre, code {
  max-height: 500px;
  overflow-y: auto;
  white-space: pre !important; 
  overflow-x: auto
}
```

```{r}
#| message: false
#| echo: false

# Packages and Source
library(tidyverse)
library(psych)
library(janitor)
library(jmv)
library(kableExtra)
library(lubridate)
```

```{r}
#| message: false
#| echo: false

# Global Settings
theme_set(theme_classic())
options(tibble.width = Inf)
options(tibble.print_max = Inf)
```

## Background

### Personal Statement

I collect moments that clarify my purpose. Among the most powerful are those shared with patients grappling with the questions my research seeks to answer. One such moment belonged to a college student struggling to determine whether an intoxicated sexual experience was coercive or consensual. Alcohol-related sexual violence is the focus of my work, and even then, I could not protect this young woman from what had already happened. Her session laid bare the blurry lines available for navigating ambiguous sexual events and their consequences.

My long-term career goal is to work as part of an interdisciplinary research team focused on promoting women's health by reducing interpersonal violence and hazardous substance use. Specifically, I want to leverage mobile technologies and the scaffolding of preexisting systems to deliver scalable, evidence-based interventions in clinical and educational settings. I am committed to disentangling complexities across the spectrum of sexual behavior, from consensual to non-consensual and sober to incapacitated via substance use, because I believe facilitating a more comprehensive understanding of these behaviors can transform the sexual violence prevention landscape.

A paradigm shift is sorely needed given the current state of the primary prevention literature and I am on my way to becoming a well-rounded psychological scientist who advances systematic change. In this vein, I have deliberately integrated my teaching, research, and clinical training experiences to prepare for a dynamic career that makes the best use of my expertise across these domains. For example, I led discussion sections on human sexuality to further my goal of using comprehensive sexuality education to prevent sexual violence among college students, then went on to complete a clinical practicum rotation at our institution's health center to inform my research focused on this population.

Notably, my path has not been without challenges. I was the first student in a new lab established during a global pandemic and then diagnosed with a rare genetic condition that causes chronic pain. My illness tested my resilience as it forced me to balance the demands of a brilliantly rigorous graduate program while fighting to be taken seriously in our healthcare system. Perseverance was critical because for me and countless others, seeking treatment means signing up to be judged and patronized with your pain dismissed. The emotional consequences of these interactions are also common among victims of sexual violence: feeling powerless, confused, and alone.

Experiencing this powerlessness firsthand clarified my purpose on a deeper level. My research aims to help stop sexual violence before it starts because I know what it feels like when your bodily autonomy is denied and will never be satisfied dealing solely with the fallout. I provide trauma-informed care because patients deserve to be seen and heard--especially by those advancing science via knowledge derived from their most vulnerable moments. I am honored to do this work and look forward to tackling new challenges in the years ahead, secure in the knowledge that with each passing milestone I get one step closer to enacting positive change for all the patients whose moments I carry with me.

### Curriculum Vitae

CV can be accessed [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/CV-7-2024.pdf).

### Transcript

Transcript can be accessed [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/unofficial-transcript-Jul-24.pdf).

## Research Experiences

### Research Statement

My research program aims to reduce sexual violence via a deeper understanding of sexual behavior following alcohol consumption. In pursuit of this goal, I integrate qualitative, quantitative, and experiencing sampling methods to investigate how processes spanning multiple ecological levels shape sexual consent dynamics and patterns of substance use, thereby heightening the risk for sexual violence.

Personal characteristics, relationships, and broader social forces shape the lens through which we interpret our experiences. The information we use to form belief systems and make decisions is filtered through this lens, which is highly consequential in the context of sexual behavior. For instance, the #MeToo movement drew the public's attention to sexual violence and changed how some adults view consent and approach sexual interactions (Ward, Nardella, Hamilton & Walsh, 2024). Awareness of #MeToo and the nature of resultant changes varied as a function of political ideology and gender, demonstrating how individual and societal factors shape sexual consent perceptions and behaviors.

Qualitative data from our #MeToo study also identified sexual victimization history and substance use as issues that complicate sexual consent, prompting subsequent efforts to examine these individual-difference factors. Histories of sexual violence victimization were indirectly related to the use of nonverbal and passive sexual consent cues through lower consent feelings among college students (Walsh, Lowe, & Ward, 2024), suggesting that mutual consideration and communication throughout sexual events may help to ensure both parties are willingly participating. Alcohol may undermine efforts to engage in this attentive, ongoing process, and college men are at high risk for both sexual violence perpetration and hazardous alcohol use, so I led a study examining how college men's alcohol-related attitudes and rape-supportive beliefs influence process-based consent. My findings indicated that positive attitudes toward alcohol-involved sexual experiences were associated with weaker process-based consent, and this negative association was exacerbated when perceived peer support of abuse was high (Ward, Hassanpour, & Walsh, under review). Collectively, these findings clarify the interplay between intrapersonal, interpersonal, and cultural risk factors that mitigate effective sexual consent.

My future research will progress the science in two important ways. Traditional quantitative approaches can miss the detailed insights needed to contextualize study findings while behavioral measures of substance use and consent often rely on retrospective self-reports. To gain a more holistic perspective of the proximal and temporal associations between sexual behavior and substance use, I will analyze qualitative and longitudinal quantitative data in my dissertation. Specifically, I will use qualitative methods to explore how young adults navigate intoxicated sexual experiences and determine for themselves whether these experiences are consensual. My lab and I also collected daily diary data from college students to investigate how substance use and subjective intoxication influence sexual risk and consent dynamics at the event level. Experience sampling methods are acceptable to people who use substances (Wyant, Moshontz, Ward, Fronk, & Curtin, 2023), and surveying students 'in the wild' can address methodological hurdles by enabling us to probe event-level associations. Collecting rich qualitative data and dynamic daily reports will strengthen my examination of intoxicated, consent-related behaviors by facilitating a more nuanced analysis of variability within and between events.

Ultimately, I aspire to leverage the strengths of multimodal ambulatory assessment and mixed methodologies to better account for the real-world complexities underlying substance-involved sexual experiences. I intend to advance understandings of the precursors and contextual factors that contribute to these incidents on college campuses and then expand the scope of my work to include members of the armed forces and individuals impacted by the criminal legal system. My goal is to inform, develop, and deliver risk-reduction and health promotion interventions aimed at preventing sexual violence and fostering healthy, consensual sexual behaviors resistant to the harmful effects of alcohol and other drugs.

### Research Products

#### Submitted Version of the FYP

1.  **Ward, S. B.**, Hassanpour, P., Walsh, K. (2023). Process-based consent among college men: The interactive influence of hostility toward women, perceived peer norms, and alcohol-involved sexual consent attitudes. *Under Review.*

Full-text PDF available [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/My%20Papers/Ward%20et%20al.%2C%20submission%20-%20sex%20roles.pdf).

**Abstract**

Process-based consent refers to the ongoing consideration of a partner's willingness to engage in sexual activity and continual communication with that partner. Drawing upon research identifying individual-, situational-, and peer-level factors that contribute to men's perpetration of sexual violence, the current study used online survey data and hierarchical linear regression to examine how permissive attitudes toward sexual consent experiences involving alcohol use, hostility toward women, and perceived peer norms in support of abuse influence process-based consent among college men (N = 463; M age = 19.32; 93% heterosexual, 64% white, non-Hispanic). Findings revealed all three primary study variables were independently associated with weaker process-based consent while adjusting for fraternity membership, relationship status, and sexual history. A significant three-way interaction indicated that the negative association between permissive alcohol-involved sexual consent attitudes and process-based consent was exacerbated when both hostility toward women and peer support of abuse were high, implying a complex relationship between multiple forces shaping college men's approaches to sexual consent. Implications for consent education and violence prevention are discussed, including the need for interventions that address process-level interactions between personal beliefs and gendered peer group norms concerning alcohol use and sexual behavior.

Relative contributions:

-   Conceptualization - 60%
-   Design - 60%
-   Analyses - 100%
-   Writing - 70%

#### Published Empirical Studies

1.  **Ward, S. B.**, Nardella, S., Hamilton, K. W., Walsh, K. (2024). "I Didn't Realize How Common It Was:" A Mixed-Methods Study Examining Changes in Perceptions of Sexual Assault, Sex and Consent, and Sexual Behavior as a Function of the #MeToo Movement. *The Journal of Sex Research*, 1--15. <https://doi.org/10.1080/00224499.2024.2352555>

Full-text PDF available [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/My%20Papers/Ward%20et%20al.%2C%202024%20-%20metoo.pdf).

**Abstract**

This mixed-methods study examined awareness and perceived legitimacy of the #MeToo movement and how #MeToo changed perceptions of sexual assault and consent, as well as sexual interactions, in the United States. Adults residing in the U.S. were recruited through CloudResearch to complete an online survey in 2021. Quantitative data from 680 participants (*M* age = 45.8, 60% women, 77.4% White) indicated moderate awareness and perceived legitimacy of the #MeToo movement; Black, LGBQ+, and more politically liberal respondents had greater #MeToo awareness while younger, more liberal respondents, and those with greater rejection of rape myths rated #MeToo more legitimate. Among 354 participants (*M* age = 45.0, 65.3% women, 76.4% White) who answered at least one open-ended question with regard to changes resulting from the #MeToo movement, thematic analysis revealed nine primary themes: (1) Describing change; (2) Change in understanding; (3) More cautious; (4) Wrong or requires reporting/punishment; (5) Gendered social norms; (6) Easier to talk about; (7) #MeToo rhetoric; (8) Consequences for survivors, and (9) Empowerment. Several sub-themes were identified among the primary themes and implications for sexual assault prevention and response were discussed. The study findings are contextualized by social constructionism, with a particular focus on gender-based power dynamics, contributing to a growing literature documenting the cultural impact of the #MeToo movement.

Relative contributions:

-   Conceptualization - 50%
-   Design - 50%
-   Analysis - 20%
-   Writing - 50%

2.  Walsh, K., Lowe, S. R., & **Ward, S.** (2024). Sexual Victimization, Emotion Dysregulation, and Sexual Consent Feelings and Communication During a Recent Sexual Encounter. *The Journal of Sex Research*, 1--9. <https://doi.org/10.1080/00224499.2024.2368044>

Full-text PDF available [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/My%20Papers/Walsh...%20Ward%2C%20et%20al.%2C%202024.pdf).

**Abstract**

Sexual consent has been a major focus of campus campaigns to reduce sexual violence (SV). However, these campaigns often educate students about consent with little attention to the complex ways consent can be experienced, expressed, and interpreted by others. Further, little research has focused on the consent feelings and communication cues of students who have a history of SV, nor have studies examined how the ability to attend to and regulate emotions relates to internal feelings or external communication of consent. This secondary analysis examined SV histories, emotion dysregulation, and internal and external consent cues in a sample of 610 college students (72% women) who completed an online self-report survey. Findings revealed that students with SV histories (n = 257) reported greater emotion dysregulation, lower internal consent, and greater use of passive external consent cues compared to students without SV histories; greater emotion dysregulation was associated with lower internal consent. In a path model, SV history was indirectly related to use of direct nonverbal and passive consent cues through greater emotion dysregulation and lower internal consent. Contrary to expectations, higher internal consent was associated with greater use of passive external consent cues. Consent programming could be augmented to encourage attention to and regulation of emotions in sexual situations to improve communication and sexual well-being.

Relative contributions:

-   Conceptualization - 20%
-   Design - 20%
-   Analysis - NA
-   Writing - 30%

#### Cross-Lab Collabs

1.  Wyant, K., Moshontz, H., **Ward, S.B.**, Fronk, G., Curtin, J. J. (2023). Acceptability of Personal Sensing Among People With Alcohol Use Disorder: Observational Study. *JMIR Mhealth Uhealth*, 11(e41833), <https://doi.org/10.2196/41833>.

Relative contributions:

-   Conceptualization - 10%
-   Design - 10%
-   Analysis - NA
-   Writing - 10%

Full-text PDF available [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/My%20Papers/Wyant...%20Ward%2C%20et%20al.%2C%202023.pdf).

#### Post-bacc Publications

1.  **Ward, S. B.**, Boger, M., Fleishman, A., Shenkel, J., Calvo, A., Pomahac, B., Zwolak, R., Krishnan, N., Richman, K., Rodrigue, J.R. (2020). Attitudes toward organ, tissue and vascularized composite allograft (VCA) donation and transplantation: A survey of United States military veterans. Transplantation. Full-text [PDF](https://github.com/sbward23/sbw_portfolio/blob/main/materials/My%20Papers/Ward%20et%20al.%2C%202020.pdf). <https://doi.org/10.1097/tp.0000000000003376>

2.  Robinson, K. A., Carroll, M., **Ward, S. B.**, Osman, S., Chhabra, K. R., Arinze, N., Amedi, A., Kaafarani, H., Smink, D., Kent, T., Musa M. Aner, Brat, G. (2020). Implementing and evaluating a multihospital standardized opioid curriculum for surgical providers. Journal of Surgical Education, 77(3), 621--626. Full-text [PDF](https://github.com/sbward23/sbw_portfolio/blob/main/materials/My%20Papers/Robinson...%20Ward%20et%20al.%2C%202019.pdf). <https://doi.org/10.1016/j.jsurg.2019.12.012>

#### Capstone Presentation

***Sexual Violence in University Students: A Multi-Level Approach to Understanding the Perpetration of Sexual Aggression***

-   Date presented: April 26, 2024
-   PDF of slide deck can be accessed [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/capstone.pdf).

#### Lunch and Learn Presentation

***Process-Based Consent Among College Men: The Interactive Influence of Hostility Toward Women, Perceived Peer Norms, and Alcohol-involved Sexual Consent Attitudes***

-   Date Presented: February 20, 2023

Sexual consent has been defined as an individual's voluntary, sober, and conscious willingness to engage in sexual activity (Willis et al., 2021). Process-based consent (PBC) builds on this definition to encourage ongoing and collaborative communication throughout sexual encounters. A potential threat to engaging in PBC is alcohol-involved sexual activity, which is highly prevalent among college students (Herbenick et al., 2010). Alcohol-related impairment can diminish an individual's ability to consent, yet perceptions of consent for alcohol-involved sexual activity vary amongst college students (Hills et al., 2021). Additional barriers to PBC may stem from hostile masculine norms (e.g., acceptance of violence toward women), as researchers have found that men who endorse these norms also report more negative attitudes toward consent (Hermann et al., 2018). The current study sought to examine the interactive effects of positive attitudes toward alcohol-involved sex (AIS) and hostility toward women (HTW) at the individual- and peer-level (i.e., perceptions of peers' HTW) on college men's endorsement of PBC. Quantitative analyses were conducted using self-report survey data from undergraduate men (N = 463) at a heavy-drinking public university in the midwestern United States. After controlling for the effects of heavy episodic drinking and relationship status, a series of linear regressions revealed a significant, negative association between positive attitudes toward AIS and PBC (β = -0.56, F(1, 459) = 191.75, p \< .001), which was negatively moderated by HTW (β = -0.19, F(1, 457) = 12.29, p \< .001) and perceptions of peers' HTW (β = -0.18, F(1, 457) = 18.50, p \< .001). Results imply that positive attitudes toward AIS are not conducive to PBC. Approval of AIS may represent a salient threat to PBC among college men who hold hostile attitudes toward women and/or among those who believe their peers hold hostile attitudes toward women. The authors conclude by discussing the implications of these findings and their relevance to campus sexual health initiatives.

#### Posters

*\*\*Mentored presentation by an undergraduate mentee.*

1.  **\*\*Does Anxiety Moderate the Effect of Peer Drinking Norms on Past-Month Alcohol-Induced Blackouts? An Observational Study Among College Students.** Date presented: April 17, 2024.

    Social anxiety and drinking for social reasons have been positively associated with alcohol consumption among college students.^1-2^ Prior work also suggests that emotion dysregulation, peer norms, and drinking for anxiety-coping can lead to alcohol-related problems.^3^  Still, it remains unclear how these factors relate to alcohol-induced blackouts (AIBs). Given the high comorbidity of social and generalized anxiety, the current study extends previous research by investigating whether generalized anxiety moderates the association between peer drinking norms and AIBs. We hypothesized that generalized anxiety symptoms and peer drinking norms would interact such that students with more anxiety would be more susceptible to higher peer drinking norms, resulting in more frequent past-month AIBs. We further hypothesized that the interaction between generalized anxiety and peer drinking norms would explain unique variance in AIBs when controlling for emotion dysregulation and other alcohol-related problems. Overdispersed data prompted two quasi-Poisson regressions. Contrary to Hypothesis 1, the results of our first model indicated that the interaction between generalized anxiety and peer drinking norms was significant but negative (β = -0.006, SE = 0.003, z = -2.207, p = .027), which suggests higher peer drinking norms increased blackout frequency and that this effect was less pronounced in students with greater anxiety. In line with our second hypothesis, the interaction remained significant after adjusting for covariates (β = -0.031, SE = 0.012,  z = -2.628, p = .009), which further suggests generalized anxiety symptoms buffer the influence of peer drinking norms on AIBs. Implications for intervention and future research are discussed.

2.  **\*\*Associations Between Lifetime Sexual Violence and Engaging in Choking During Sex: Findings from an Undergraduate Sample.** \
    Date presented: November 15, 2023.

    Background: Sexual choking is a behavior that some engage in to heighten sexual pleasure, replicate gendered sex scripts, and explore kinky behaviors (Herbenick, 2020; Tarr, 2016). However, choking can lead to unintended and dangerous consequences because applying external pressure on the neck can restrict airflow and cause brain anoxia (Sauvageau & Boghossian, 2010; Hou et al., 2023). Women are more likely to be choked than men and men are more likely to choke a partner compared to women (Bridges et al., 2016; Wright et al., 2015). Lesbian, gay, and bisexual students are more likely to engage in choking compared to straight students (Herbenick et al., 2021). Although likelihood of engaging in BDSM is not associated with experiencing sexual trauma (Ten Brink, 2021), engaging in kink play or rough sex may be empowering for survivors of sexual trauma (Cascalheira, 2021). However, to our knowledge, associations between sexual violence and sexual choking (not in the context of a violent encounter) have yet to be explored. Methods: The current study seeks to understand the relationship between sexual violence and choking during sex. College students (N = 228; 56.1% female; 69.7% white; 86.4% heterosexual; mean age = 18.6) completed an online survey about sexual behavior including questions about the frequency of engaging in choking, the degree to which they enjoyed it, and how often they engaged in choking despite not enjoying it. Results: Findings indicated that 56.6% had participated in choking at least once, 31% had engaged in choking more than 1x/month, and 16% indicated that they engaged in choking despite not enjoying it. Consistent with prior work, women were more likely to report choking compared to men, Χ2 = 6.23, p = .01, but contrary to expectations, no differences were observed for sexual orientation, Χ2 = 1.12, p = .29. 44.7% of the sample reported lifetime sexual violence. After controlling for gender, sexual orientation, and age of first consensual sexual encounter, participants who reported sexual victimization during their lifetime had increased odds of reporting engagement in choking (OR: 1.962, 95% CI: 1.139-3.378) and choking despite not enjoying it (OR: 5.501, 95% CI: 1.994-15.178). Conclusion: Future qualitative research could explore the ways in which individuals who engage in choking conceptualize their own sexual agency and desires along with their willingness and ability to communicate consent to engage in choking during sex.

3.  **\*\*Manner of Drinking as a Mediator of Emotional Impulsivity and Alcohol-Related Problems.**\
    Date presented: April 19, 2023.

4.  **Alcohol-Induced Blackouts and Nonconsensual Sexual Experiences Among Undergraduate Men.** November 18, 2022.

    College students who have previously engaged in sexually coercive behavior tend to have significantly higher sex-related alcohol expectancies, while victims of unwanted sexual contact often report higher alcohol consumption, use fewer protective strategies, and experience more negative consequences due to their alcohol use (Palmer et al., 2010). One such consequence is experiencing more alcohol-induced blackouts (AIBs), or periods of memory loss for all or part of a drinking episode, and these periods of memory loss have been associated with a wide range of negative outcomes including sexual assault (SA) (Haas et al., 2016; Voloshyna et al., 2018). Individuals who have experienced SA victimization may use increasing amounts of alcohol to self-medicate assault-related distress (Stappenbeck et al., 2022) and those with a history of SA perpetration may more frequently drink to the point of blacking out as a tactic to avoid culpability in the pursuit of an intoxicated sexual partner (Ven & Beck, 2009) yet findings on what could be reciprocal associations between SA and AIBs have been mixed (Jaffe et al., 2022). Most of these associations have been observed among college women despite college men being more likely to report intentions to blackout (Wombacher et al., 2019), which is understandable considering the higher risk for SA victimization among women. However, given that men who have experienced or perpetrated  (Neal & Fromme, 2007) sexual violence often report heavy alcohol use, research examining whether victimization or perpetration histories are associated with AIBs among college men is needed. The current study characterized the prevalence and sociodemographic correlates of past-month AIBs among undergraduate men and examined whether nonconsensual sexual experiences since age 14 were significant predictors of AIBs within this population. Among 473 undergraduate men (40% fraternity members; mean age = 19.3; 37% students of color) who completed an online survey assessing alcohol use and sexual history, 61.73% reported experiencing an AIB within the past month. Past-month AIB severity was positively associated with fraternity membership (*β* = 0.67, F(1,471) = 41.21, *p* \< .001)  and white racial identity (*β* = -0.24, F(1,471) = 4.66, *p* = .03). More than one-third of the sample reported at least one nonconsensual sexual experience since age 14 (39.74% SA victimization, 8.48% perpetration) and a series of quasi-Poisson regressions revealed that both SA victimization (*β* = 0.44, F(1,461) =17.31, *p* \< .001) and perpetration (*β* = 0.40, F(1,470) = 6.04, *p* = .01) were significant predictors of past-month AIB severity. Alcohol and sex education programs are often conducted separately, and these data suggest it may be important to consider their reciprocal associations. Campus health centers might consider screening for SA victimization among college men with problematic alcohol use, as our data suggest victimization experiences may have previously been underestimated within this population (Fedina et al., 2018; Mellins et al., 2017) already at increased risk of incurring alcohol-related harm.

5.  **Sexual Consent Among College Men: The Moderating Effects of Alcohol-Involved Sex and Hostility Toward Women.** \
    Date presented: November 5, 2022.\
    \
    Sexual consent has been defined as an individual's voluntary, sober, and conscious willingness to engage in sexual activity (Willis et al., 2021). Process-based consent (PBC) builds on this definition to encourage ongoing and collaborative communication throughout sexual encounters. A potential threat to engaging in PBC is alcohol-involved sexual activity, which is highly prevalent among college students (Herbenick et al., 2010). Alcohol-related impairment can diminish an individual's ability to consent, yet perceptions of consent for alcohol-involved sexual activity vary amongst college students (Hills et al., 2021). Additional barriers to PBC may stem from hostile masculine norms (e.g., acceptance of violence toward women), as researchers have found that men who endorse these norms also report more negative attitudes toward consent (Hermann et al., 2018). The current study sought to examine the interactive effects of positive attitudes toward alcohol-involved sex (AIS) and hostility toward women (HTW) at the individual- and peer-level (i.e., perceptions of peers' HTW) on college men's endorsement of PBC. Quantitative analyses were conducted using self-report survey data from undergraduate men (N = 463) at a heavy-drinking public university in the midwestern United States. After controlling for the effects of heavy episodic drinking and relationship status, a series of linear regressions revealed a significant, negative association between positive attitudes toward AIS and PBC (β = -0.56, F(1, 459) = 191.75, p \< .001), which was negatively moderated by HTW (β = -0.19, F(1, 457) = 12.29, p \< .001) and perceptions of peers' HTW (β = -0.18, F(1, 457) = 18.50, p \< .001). Results imply that positive attitudes toward AIS are not conducive to PBC. Approval of AIS may represent a salient threat to PBC among college men who hold hostile attitudes toward women and/or among those who believe their peers hold hostile attitudes toward women. The authors conclude by discussing the implications of these findings and their relevance to campus sexual health initiatives.

6.  **Associations between Sexual Assault History and Sexual Consent Feelings and Behaviors among College Students: A Path Model Involving Alcohol Use Prior to Sex & Emotion Dysregulation.** \
    Date presented: November 18, 2021.

    Sexual assault is a significant problem among college students with approximately one in four college women reporting some form of unwanted sexual contact. Although sexual assault and substance use have been linked with emotion dysregulation and sexual risk taking, no studies of which we are aware have examined associations between sexual assault, substance use, emotion dysregulation and sexual consent feelings and behaviors. The current study examined associations between sexual assault history, substance use prior to sex, emotion dysregulation, and sexual consent feelings (i.e., willingness) and behaviors (i.e., consent communication) during a recent sexual encounter. Among 610 college students (72% female; mean age = 21.3; 44% underrepresented racial/ethnic groups) who completed an online survey, approximately 31% reported unwanted sexual contact since age 14 (15.7% men, 35.3% women), and 13.9% (15.9% men, 13.4% women) reported substance use prior to their most recent sexual encounter. The best fitting path model (*χ2* (*df* = 3) = 5.25, *p* = .15, *CFI* = .98, *RMSEA* = .03, *SRMR* = .02) revealed that a history of sexual assault was significantly positively associated with overall emotion dysregulation, which in turn was positively associated with substance use prior to sex but negatively associated with internal willingness during their most recent sexual encounter. Higher levels of willingness, in turn, were positively associated with the use of more external consent communication behaviors. Findings suggest that consent education programs for college students may need to consider the role of sexual assault victimization history and emotion dysregulation in addition to alcohol as factors that impact willingness to engage in sexual activity and downstream communication about consent.

7.  **\*\*Sexual Assault Resistance Strategies, Blame, and Mental Health Outcomes.** \
    Date presented: November 4, 2021.

    Sexual assault has been associated with posttraumatic stress disorder (PTSD) and depression. The resistance strategies a survivor employs (forceful, non-forceful, or passive) and the blame one attributes (towards self or rapist) for an assault may affect these outcomes. To our knowledge, these factors have not been analyzed together. We hypothesized that greater use of passive resistance strategies and higher levels of self-blame would be positively associated with PTSD and depressive symptoms. We used online survey data from 62 male and female university students who reported an unwanted sexual experience. As predicted, bivariate correlations showed that greater use of passive resistance strategies was positively correlated with PTSD (*r*=.51, *p*\>.001) and depression (*r*=.27, *p*\>.05) symptoms, and higher levels of self-blame (*r*=.25, *p*\>.05). Higher levels of self-blame were also positively associated with PTSD (*r*=.40, *p*\>.01) and depression (*r*=.31, *p*\>.05) symptoms. The finding that passive resistance strategies were associated with negative mental health outcomes may reflect internalization of rape myths, whereby survivors who do not actively fight back or clearly say no consider themselves responsible for their assaults. Findings also may inform treatment of patients with sexual victimization histories by targeting the role of self-blame in the onset of PTSD and depression.

8.  **Sexual Assault Victimization and Perpetration among those Involved in Greek Life and Varsity Athletics: A Large National Study.** \
    Date presented: November 19, 2020.

    Studies on whether college students involved in fraternities, sororities, and campus athletics are at increased risk for sexual assault victimization and perpetration have yielded mixed findings, and many existing studies are dated and focus on samples at single universities. Past research has found that fraternity and sorority members consume more alcohol than nonmembers and suggest this as a potential factor contributing to higher rates of sexual assault among Greek members (Franklin, 2016; Murnen & Kohlman, 2007). Fraternity and sorority members are also more likely to report using drinking protective behavioral strategies (PBS), a trend mirrored by varsity athletes (citation?). PBS have been suggested as a potential preventive method of reducing sexual assault risk (Neilson et al., 2018). The current study seeks to examine these patterns using a large national sample drawn across multiple campuses between 2011-2013. Participants were 88,975 students (M age = 22.3, SD = 6.2; 72.7% White; 65.2% female, 32.5% male, 0.3% transgender) from 62 different campuses. Students received a survey broadly about college student health. Approximately 11% (12.1% women, 10.6% men) of participants reported involvement in a fraternity or sorority and 8.0% (7.5% women, 9.2% men) reported involvement in varsity athletics in the previous 12 months. Chi-square analyses run separately by gender revealed that female, male, and transgender students involved in Greek life were significantly more likely than students who were not to report unwanted sexual touching (ps \<.001), attempted sexual penetration (ps \<.01), and completed penetration (ps \<.05). Similar patterns emerged for female, male, and transgender students involved in varsity athletics. Among students who reported any drinking, students of all genders involved in Greek life and athletics were more likely to report that someone had sex with them without their consent while they were drinking (ps \<.01), and male and transgender students involved in Greek life (ps \<.001) and varsity athletics (ps \<.05) were more likely to report that they had sex with someone without their consent while drinking. Notably, protective behavioral strategies do not seem to serve as a protective factor for sexual assault among these populations. Findings have important implications for collegiate athletics as well as Greek organizations, suggesting the alcohol education programs in use with these students across campuses need to incorporate curriculum addressing issues around consent. Further, there is a heightened need for evidence-based sexual violence prevention programming for sorority women, fraternity men, and varsity athletes that covers substance-facilitated sexual assault.

#### Workshops

**Attended**

-   Academic Job Materials Workshop - Princeton Psychology Postdoctoral DEI Committee
-   Mixture Modeling and Latent Class Analysis -- Daniel J. Bauer, PhD, and Doug Steinley, PhD
-   Responsible Authorship and Publication -- Margarita Kaushanskaya, PhD,
-   Research Misconduct and Policies for Handling Misconduct -- Brain Fox, PhD
-   Ins and Outs of F and K Awards -- Waisman Center PIs and Grants Team
-   Data Acquisition, Analysis, Sharing, and Ownership -- Brian Christian, PhD
-   Collaborative Research -- Doug Dean III, PhD
-   Mentor and Mentee Responsibilities and Relationships -- Rebecca Alper, PhD, CCP‐SLP
-   Peer Reviews and NIH Grant Reviews -- Audra Sterling, PhD
-   Conflict of interest---Personal, Professional, and Financial -- Maureen Durkin, PhD, DrPH
-   Ethical Conduct of Human and Animal Subjects Research -- Katherine Hustad, PhD, CCC‐SLP
-   Secure and Ethical Data Use and Management -- Brittany Travers, PhD
-   Introduction to Structural Equation Modeling -- Daniel J. Bauer, PhD, and Patrick J. Curran, PhD

**Focused on DEI Practices and Principles**

-   Breaking the Bias Habit -- William T.L. Cox, PhD
-   Promoting Safe and Inclusive Research Environments -- Leann DaWalt, PhD
-   The Scientist as a Responsible Member of Society -- Karla Ausderau, PhD
-   Change in Implicit Bias -- Calvin Lai, PhD
-   Cross-Race Friendship -- Elizabeth Page-Gould, PhD
-   Exploring 'the story' of inequality: Scientific investigations into the antecedents, experience, and reduction of group-based inequality -- Demis Glasford, PhD
-   Inequalities in Young Adult Criminal Justice Contacts: Patterns and Collateral Consequences for Parents and Siblings -- Courtney Boen, PhD, MPH

## Clinical Experiences

### Theoretical Orientation Statement

My theoretical orientation is rooted in the cognitive-behavioral tradition. Third-wave principles guide my approach to case conceptualization and intervention, which strives to be holistic in nature by considering the context of an individual's experience. The cognitive-behavioral foundation, core concepts, and process-based focus of the third wave grounds my practice in science while enabling integration and flexibility.

Classic cognitive behavioral therapy (CBT) focuses on patterns of problematic thinking and their impact on mood and behavior. I use CBT and interventions derived from this tradition in my practice, such as CBT-E for eating disorders and cognitive processing therapy for PTSD. In the provision of CBT, I help my patients examine the content of automatic thoughts and their perceptions of what prompted these thoughts to facilitate a more balanced perspective. Third-wave principles influence my CBT service provision in that I still attend to the content of thoughts but shift more focus to their function and how my patients relate to them. Both approaches are valuable and research identifying moderators of treatment response can help to inform decisions about when to use which method and with whom.

Third-wave methods highlight the role of mindfulness, acceptance, emotions, values, and meta‐cognition. Each of these constructs has an extensive knowledge base that connects their role in treatment to evidence-based mechanisms of change. I am partial to third-wave therapies including acceptance and commitment therapy (ACT) and dialectical behavior therapy (DBT) because of their focus on these pathways to change and utility in guiding case conceptualization. Specifically, I use a bottom-up approach to conceptualization that considers a patient's symptoms in the context of their goals, history, culture, relationships, and strengths, as well as the domains they identify as most critically impacted. Doing so prevents me from using their symptoms to make a premature diagnosis and then subconsciously building a conceptualization and treatment plan around that disorder. This is a strategy I aim to avoid given the high rate of comorbidity among diagnostic categories originally theorized to have distinct etiologies, and because fixating on so-called disorders means narrowly focusing on deficiencies and dysfunction rather than health and wellness.

In contrast, my holistic approach to conceptualization allows me to identify transdiagnostic processes of change consistent with a patient's presentation, motivation for therapy, and the dimensional nature of psychopathology. Centering core processes also means conceptualizations and treatment plans can be more adeptly tailored as my understanding of a patient grows and their needs evolve throughout therapy. However, flexibility does not mean I readily abandon evidence-based protocols whenever a patient wants to switch gears. I strive to balance flexibility and fidelity in my practice, integrating complementary approaches to leverage the strength of compatible strategies. I find this most useful when adaptations increase cultural sensitivity or serve to benefit the working alliance in a meaningful way.

### Clinical Training Record

```{r,echo=FALSE}
#| label: tbl-all
#| tbl-cap: "Overall Hours"

# create data frame
all_summ <- data.frame(
  Summary = c("Intervention", "Assessment", "Total Direct", "Supervision"),
  Hours = c("328.75", "93.50", "422.25", "244.75"))
# format table
kable(all_summ, booktabs = TRUE, align = 'lr', col.names = c("Summary", "Hours")) %>%
  kable_styling(full_width = FALSE)
```

#### Practicum Experiences

**University Health Services, University of Wisconsin-Madison**

September 1, 2022 - May 13, 2023, 16 - 20 hours per week

-   Individual Counseling: 5 -- 8 hours/week.
-   Group Counseling: 2 hours/week.
-   Individual Supervision: 1.5 hours/week.
-   Group Supervision: 1.5 hours/week.
-   In-service Training: 1.5 hours every other Friday.
-   Group Therapy Seminar/Supervision: 1.5 hours every other Friday.
-   Administrative/Case Notes: 3-5 hours/week.

As a practicum trainee in the Mental Health Services division of our institution's health center, I maintained a clinical caseload of 10-18 young adults at a time, treating 35 individual patients in total. My patients were undergraduate and graduate students who presented with symptoms consistent with social and generalized anxiety, posttraumatic stress, attention-deficit/hyperactivity disorder, and major depressive disorder, though concerns specific to substance misuse, suicidal thoughts, self-injurious behaviors, gender dysphoria, poor body image and low self-worth were also quite common. During this clinical practicum, I gained experience working as a multidisciplinary team member within a coordinated care model. I collaborated with clinicians from other service divisions including crisis stabilization, primary care, OB/GYN, and psychiatry to provide holistic care to each patient.

```{r,echo=FALSE}
#| label: tbl-uhs
#| tbl-cap: "Hours by Type at University Health Services"

# create data frame
uhs_summ <- data.frame(
  Summary = c("Intervention", "Assessment", "Total Direct", "Supervision"),
  Hours = c("153.00", "27.00", "180.00", "87.25"))
# format table
kable(uhs_summ, booktabs = TRUE, align = 'lr', col.names = c("Summary", "Hours")) %>%
  kable_styling(full_width = FALSE)
```

**Psychology Research and Training Clinic, University of Wisconsin-Madison**

September 1, 2020 - Present

Through practicum training in the PRTC, I have provided individual therapy, conducted diagnostic and neuropsychological assessment batteries, and screened previously incarcerated individuals for psychosis, suicidality, and PTSD. In Fall of 2023, I acquired deeper theoretical knowledge and new clinical skills in the assessment and treatment of PTSD and trauma-related conditions. Elements of this internal practicum required mastering the Clinician-Administered PTSD Scale for DSM-5 (CAPS-5) and delivering closely supervised Cognitive Processing Therapy.

```{r,echo=FALSE}
#| label: tbl-prtc
#| tbl-cap: "Hours by Type at PRTC"

# create data frame
prtc_summ <- data.frame(
  Summary = c("Intervention", "Assessment", "Total Direct", "Supervision"),
  Hours = c("173.75", "68.50", "242.25", "157.50"))
# format table
kable(prtc_summ, booktabs = TRUE, align = 'lr', col.names = c("Summary", "Hours")) %>%
  kable_styling(full_width = FALSE)
```

A time2track summary of patient demographics and hours can be accessed [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/Combined%20Summary%20_%20Time2Track.pdf).

#### Integrative Assessment Reports

Five de-identified, integrated assessment reports are accessible in Box.

## Diversity, Equity, and Inclusion

### Diversity Statement

Systemic issues that threaten diversity, equity, and inclusion are inextricably linked to the roles I inhabit as a feminist research scientist and aspiring clinical psychologist.

I study sexual violence, the burden of which is inequitably distributed. Racialized and other minoritized groups are disproportionately impacted by sexual violence and the risk of victimization is compounded among those with multiple marginalized identities. I promote equity through sexual violence research by investigating the nature of these disparities and facilitating efforts to eliminate them. Rigorous science in this area of study requires a deliberate awareness of implicit biases and racist, patriarchal power structures, which shape the processes through which both disparities and scientific knowledge are produced.

The assumption that humans can be wholly objective ignores history and context, so promoting equity through research means challenging the assertion that a universal truth underlying complex psychosocial experience can be unearthed through objective scientific means (Rodriguez-Seijas et al., 2023). To this end, I subscribe to the contemporary constructivist philosophy of science and use theory aligned with its core assumptions to guide my research. For example, a constructionist framework contradicts the positivist notion of a mind-independent, absolute truth, instead positing that reality is malleable and "constructed" through social processes (DeLamater & Hyde, 1998). This framework assumes social problems are actively made, not simply revealed, a core tenet of feminist science and the scholarly literature on societal norms that perpetuate sexual and gender-based violence.

Social problems are actively made in society at large but also within our communities and institutions. Promoting diversity, equity, and inclusion therefore requires turning inward to examine the principles and practices used in the conduct of our research. There are current practices in psychological science that exclude scholars from marginalized backgrounds, which perpetuates white supremacy (Buchanan et al., 2021). I use open science practices that facilitate sharing and reuse of research assets to help mitigate inequalities disadvantaging under-resourced scholars via increased accessibility. The open science movement explicitly features collaborative, prosocial ideals known to engage members of under-represented groups and cultivate a more diverse culture (Murphy et al., 2020). These ideals reflect my values and align with my goal of working as part of a cooperative, justice-oriented research team.

Finally, it is paramount that I consider how systemic oppression and prejudice give rise to inequitable mental health outcomes. I have been fortunate to receive specialized training to enhance my cultural competence and have learned about how oppressive forces can impact mental health directly from those affected. My work with patients whose backgrounds differ from my own has also highlighted the value of diverse perspectives and many ways in which culture and spirituality can promote human flourishing. It is a blessing to connect with people from varied communities and I appreciate the responsibility that comes with being trusted by someone who society has forced to remain guarded.

### DEI Training Record

#### Case Conceptualization

An integrative assessment report, including case conceptualization and treatment plan, for a patient intersecting marginalized identities is accessible in Box.

#### DEI - Workshops, Seminars, and Lectures

**Clinical Service Provision**

-   International Student-Clients: Clinical Considerations -- Wei-Chiao Hsu 徐薇喬, PhD, LPC
-   Recognizing and Addressing Bicultural Identity Conflict -- Geetanjali Deole, MS, LPC

**Research, cross-listed**

-   Breaking the Bias Habit -- William T.L. Cox, PhD
-   Promoting Safe and Inclusive Research Environments -- Leann DaWalt, PhD
-   The Scientist as a Responsible Member of Society -- Karla Ausderau, PhD
-   Change in Implicit Bias -- Calvin Lai, PhD
-   Cross-Race Friendship -- Elizabeth Page-Gould, PhD
-   Exploring 'the story' of inequality: Scientific investigations into the antecedents, experience, and reduction of group-based inequality -- Demis Glasford, PhD
-   Inequalities in Young Adult Criminal Justice Contacts: Patterns and Collateral Consequences for Parents and Siblings -- Courtney Boen, PhD, MPH

#### DEI - Service

-   **Committee for Accommodations:** Collaborates with departmental faculty and graduate-student peers on initiatives designed to increase the accessibility of disability accommodations among psychology students and facilitate management of accommodations requests among instructors and teaching assistants

-   **Diversity, Equity, and Inclusion (DEI) Panel:** Served on the DEI panel at Recruitment Weekend

#### Psychology Research Experience Program (PREP) Mentor

During the summer of 2021, I had the opportunity to mentor a promising undergraduate student, Rachna Iyer, through the PREP program at UW-Madison. This program, supported by the National Science Foundation, focuses on providing students from underrepresented backgrounds with intensive research experiences in psychology and neuroscience.

In my role as a mentor, I introduced Rachna to the sexual violence literature pertaining to posttraumatic stress, providing her with a foundational understanding necessary for conducting meaningful research in her area of interest. Over the course of the program, I guided her through the entire scientific process---from the initial stages of identifying a research question to the final stages of writing up the results.

One of my primary goals was to equip Rachna with practical skills in data analysis and management. I facilitated her learning of analytic strategies using R and RStudio. This training not only enhanced her technical skills but also her confidence in handling complex datasets.

Additionally, I supported Rachna in her professional development by helping her prepare for presentations. She successfully presented her project at the 2021 PREP Symposium and the International Society for Traumatic Stress Studies (ISTSS) conference, opportunities that allowed her to engage with the broader scientific community and gain valuable feedback.

This mentorship experience was deeply enriching, as it allowed me to contribute to the growth of an aspiring clinical scientist from an under-represented background in academic psychology. After the PREP program, Rachna graduated early from the University of Michigan and then worked for Dr. Nicole Weiss, an Associate Professor and Clinical Psychologist at the University of Rhode Island, whose research program focuses on the co-occurrence of PTSD and substance use disorders.

## Teaching and Mentorship

### Teaching Experiences

Starting in the Spring semester of 2022, I served as Dr. Janet Hyde's Head Teaching Assistant for the Psychology of Human Sexuality course (PSYC/SOC 453). I led four weekly discussion sections, each consisting of 15-20 undergraduates, encouraging critical thinking and deeper engagement with course materials. My responsibilities extended to coordinating testing, managing section changes, and overseeing the grading process for the 300-student lecture course. I also delivered two guest lectures on love and gender, which were well-received by students.

During the Fall semester of 2022, I was a Teaching Assistant for the Introductory Psychology course (PSYC 202), taught by Dr. Rebecca Addington. This large, 400-student lecture course required meticulous organization and pedagogical skills; I was responsible for grading approximately 400 short essay assignments each week. Additionally, I managed logistical aspects of the course such as scheduling, proctoring, and hand-scoring makeup exams. I also supported students through tutoring during office hours, by appointment, and via email, ensuring their academic and personal support.

In January, 2024, I delivered a Guest Lecture at the School of Pharmacy for Karen Kopacek, MS, RPh. The lecture focused on emotional stress and compassion fatigue and was psychoeducational in nature, as I integrated practical skills such as mindfulness, distress tolerance, and emotion regulation to support future healthcare professionals. These experiences have collectively shaped my teaching philosophy and approach, reinforcing my commitment to fostering an interactive and supportive educational environment.

### TA Evaluations

-   Section 301: Human Sexuality, aggregate results linked [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/TA%20Course%20Evals/301_course_eval.pdf).

-   Section 302: Human Sexuality, aggregate results linked [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/TA%20Course%20Evals/302_course_eval.pdf).

-   Section 303: Human Sexuality, aggregate results linked [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/TA%20Course%20Evals/303_course_eval.pdf).

-   Section 316: Human Sexuality, aggregate results linked [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/TA%20Course%20Evals/316_course_eval.pdf).

-   PSY/SOC 453: Human Sexuality, faculty eval of TA linked [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/TA%20Course%20Evals/Faculty-eval-TA.pdf).

-   PSY202: Intro Psychology, course eval linked [here](https://github.com/sbward23/sbw_portfolio/blob/main/materials/TA%20Course%20Evals/eval_aggregate_PSY202_F22.pdf) - no instruction as TA.

### Mentorship Experiences

Since August 2020, I have been actively involved in mentorship in our Preventing Interpersonal Violence & Overcoming Trauma (PIVOT) Lab. I have provided mentorship to a diverse group of lab members, ranging from undergraduates to post-baccalaureate and master's-level research assistants.

Additionally, I have led workshops aimed at enhancing research proficiency, covering topics such as effective literature search strategies and the use of Zotero for organizing and citing scholarly articles. These workshops are designed to enhance lab members' practical, hands-on experience and are tailored to meet the evolving needs of our research program.

A significant portion of my mentorship experience has also been devoted to guiding our lab members through the graduate application process. I provide individualized support to help each person identify and apply to graduate programs and paid training opportunities that match their career aspirations in research and clinical practice. This guidance is customized to reflect the unique goals and backgrounds of each mentee, ensuring that they receive the most relevant and impactful advice.
